---
title: "E1 memory"
author: "R. Cassano-Coleman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # version 2.0.0
library(magrittr) # version 2.0.3

library(brms) # version 2.22.0
library(bayestestR) # version 0.15.0
library(emmeans) # version 1.10.0
```

This notebook analyzes memory using Bayesian binomial generalized linear mixed effects models (GLMMs).


# Set up

```{r}
set.seed(15000)
```

```{r}
data <- read_csv('../data/memory.csv')
```

Convert variables to factors.
```{r}
data %<>%
  mutate(exp_subject_id = as.factor(exp_subject_id),
         response = ifelse(response == 'Correct', TRUE, FALSE),
         scramble = factor(scramble, levels = c('8B', '2B', '1B')),
         Musician = factor(Musician, levels = c('Yes', 'No'))) %>%
  filter(!is.na(response))
```

Set the contrast for condition.

```{r}
contrasts(data$scramble) <- contr.treatment(3)
print(contrasts(data$scramble))
```

\newpage

# Main analysis

## Priors

Priors are expressed in log(odds) space.

**Intercept:** Given that chance is 50%, we assume that participants will perform somewhere between chance and ceiling. We expect the center of the distribution of accuracy to be somewhere around 75% or 80%. If we use a center of 80% and an SD of 1, 95% of the values fall between 35.1% and 96.7%.

```{r}
prior_intercept <- set_prior('normal(log(0.8 / (1 - 0.8)), 1)', class = 'Intercept')
```

**Group:** We might expect musicians to do slightly better than non-musicians, on average.

In this range, a difference in 0.25 log odds gives us about a 5% decrease in accuracy.
```{r}
prior_mus <- set_prior('normal(-0.25, 1)', coef = 'Musician1')
```

**Scramble:** We expect performance to improve as scramble level decreases. If we code 8B as reference level, then we expect 8B > 2B and 8B > 1B.

Since we're keeping the musician slope at SD = 1, we'll keep these (and the interactions) at SD = 1. This seems to be a pretty weak prior.

```{r}
prior_scramble2B <- set_prior('normal(-0.1, 1)', coef = 'scramble2')
prior_scramble1B <- set_prior('normal(-0.2, 1)', coef = 'scramble3')
```

**Interaction:** We expect no interaction between group and scramble.

```{r}
prior_int2B <- set_prior('normal(0, 1)', coef = 'Musician1:scramble2')
prior_int1B <- set_prior('normal(0, 1)', coef = 'Musician1:scramble3') 
```

**Random slope for subjects:**
*Leave this as default for now, may update.*

\newpage

## Main model with group and condition

```{r}
mus_scram <- brm(response ~ Musician + scramble + (1 | exp_subject_id), data = data,
                 family = bernoulli(),
                 prior = c(prior_intercept, prior_mus, 
                           prior_scramble2B, prior_scramble1B),
                 save_pars = save_pars(all = TRUE), iter = 5000,
                 file = '../models/E1_mus_scram')
```

```{r}
plot(mus_scram)
```

```{r}
print(summary(mus_scram), digits = 4)
```

```{r}
emm_mus_scram_s <- emmeans(mus_scram, specs = "scramble")
summary(emm_mus_scram_s)
```

```{r}
contrast(emm_mus_scram_s, method = "pairwise")
```

```{r}
emm_mus_scram_ms <- emmeans(mus_scram, specs = c("Musician", "scramble"))
summary(emm_mus_scram_ms)
```

```{r}
contrast(emm_mus_scram_ms, method = "pairwise")
```

\newpage

## Main effects

```{r}
main_BF <- describe_posterior(mus_scram,
                              estimate = "median", dispersion = TRUE,
                              ci = .95, ci_method = "HDI",
                              test = c("bayes_factor"))
print(main_BF, digits = 4)
```

Moderate evidence for a main effect of group.

\newpage

To get the main effect of scramble level, fit the "null" model with group only to compare.

```{r}
mus_only <- brm(response ~ Musician + (1 | exp_subject_id), data = data,
                 family = bernoulli(),
                 prior = c(prior_intercept, prior_mus),
                 save_pars = save_pars(all = TRUE), iter = 5000,
                 file = '../models/E1_mus_only')
```

```{r}
plot(mus_only)
```

```{r}
print(summary(mus_only), digits = 4)
```

```{r}
BF_scramble <- bayes_factor(mus_scram, mus_only)
print(BF_scramble)
```

Very strong evidence for a main effect of scramble condition.

\newpage

## Interaction between group and condition?

Add an interaction between group and condition, and compare the model with the interaction to the one without.

```{r}
mus_scram_int <- brm(response ~ Musician*scramble + (1 | exp_subject_id), data = data,
                 family = bernoulli(),
                 prior = c(prior_intercept, prior_mus, 
                           prior_scramble2B, prior_scramble1B,
                           prior_int2B, prior_int1B),
                 save_pars = save_pars(all = TRUE), iter = 5000,
                 file = '../models/E1_mus_scram_int')
```

```{r}
plot(mus_scram_int)
```

```{r}
print(summary(mus_scram_int), digits = 4)
```


```{r}
BF_int <- bayes_factor(mus_scram_int, mus_scram)
print(BF_int)
```
Moderate evidence against an interaction between group and condition.

\newpage

## Figure 2A

Visualize posterior distributions on the scale of accuracy.

Visualize with posterior estimates and 95% CrI 

```{r}
posterior_est <- as.data.frame(emm_log_rate)
```

```{r}
ggplot() +
  geom_col(aes(x = scramble, y = exp(emmean), fill = Musician), data = posterior_est,
           position = "dodge") +
  geom_errorbar(aes(x = scramble, ymin = exp(lower.HPD), ymax = exp(upper.HPD), fill = Musician), 
                data = posterior_est, position = position_dodge(width = 0.9), width = 0.2) +
  geom_point(aes(x = scramble, y = mean_response_rate, fill = Musician), data = data, 
             position = position_jitterdodge(dodge.width = 0.9, jitter.width = 0.1), alpha = 0.1) +
  theme_gray(base_size = 16) +
  scale_x_discrete(limits = rev) +
  xlab('Scramble Level') +
  ylab('Response Rate (per Minute)') +
  scale_fill_discrete(name="", labels=c('Musicians', 'Non-musicians')) +
  theme(legend.text = element_text(size = 12))

ggsave('../figures/Fig3_rate.png', width = 7, height = 5)
```



\newpage

# Years of experience 

## Priors

## Figure S1A

