---
title: "Alignment"
author: "Riesa Y. Cassano-Coleman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r libraries}
library(tidyverse) # version 2.0.0
library(magrittr) # version 2.0.3
library(nparLD) # version 2.2
library(rstatix) # version 0.7.2
select <- dplyr::select # make sure we're using tidyverse's version of select...
```

Load the data.
```{r}
data <- read_csv('../data/alignment_original.csv', show_col_types = FALSE)
```

Pivot the data longer.
```{r}
data %<>% pivot_longer(cols = -c(Musician, sub, scramble), 
                       names_to = 'level', values_to = 'value')
```

Make sure non-musicians and musicians are labelled with different numbers.
```{r}
data %<>% mutate(sub = ifelse(Musician == 'Yes', sub, sub + 45))
```

```{r}
data %<>% mutate(level = factor(level, levels = c(1,2,3,4,5,8,16), ordered = TRUE))
```

For comparisons across levels, look at nested structure only (levels 2, 4, 8, 16).
```{r}
data_nested <- data %>%
  filter(!level %in% c(1,3,5))
```


\newpage
# Alignment above chance (all levels, all conditions)

At each level, which values are above chance?
```{r}
stat.test <- data %>%
  group_by(scramble, level) %>%
  t_test(value ~ 1) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance() %>%
  arrange(scramble, level)
```

```{r}
print(stat.test, n = nrow(stat.test))
```

(Tables 2b, 3b, 4b, and 5b)


\newpage
# Three-way non-parametric ANOVA-type test
```{r}
attach(data_nested)
# scramble is a within-subject factor ("time1") - all subjects hear all conditions
# level is the other within-subject factor ("time2") - all levels are analyzed
# expertise is the between-subject factor ("group") - subjects are either musicians or non-musicians
f1.ld.f2(value, time1=scramble, time2=level, group=Musician, subject=sub,
         time1.name="condition", time2.name="level", group.name="expertise", 
         description=FALSE)$ANOVA.test
```

* No three-way interaction between scramble condition, level, and expertise group: F(6.6) = 1.70, p = .11
* Main effect of condition: F(2.8) = 5.47, p = .0012
* Main effect of level: F(2.2) = 12.14, p < .001
* Interaction between condition and level: F(6.6) = 6.67, p < .001)
* No main effect of expertise: F(1) = .443, p = .51


\newpage
# Post-hoc tests: comparison of levels (1B)

```{r}
data_nested_1B <- filter(data_nested, scramble == '1B')
```

```{r}
attach(data_nested_1B)
f1.ld.f1(value, time=level, group=Musician, subject=sub,
         time.name="level", group.name="expertise", 
         description=FALSE)$ANOVA.test
```

Main effect of level: F(2.3) = 6.69, p < .001

\newpage
```{r}
pairwise.wilcox.test(data_nested_1B$value, data_nested_1B$level,
                     p.adjust.method = "bonferroni", paired=TRUE)
```
No significant differences between levels (Table S2a).


\newpage
# Post-hoc tests: comparison of levels (2B)

```{r}
data_nested_2B <- filter(data_nested, scramble == '2B')
```

```{r}
attach(data_nested_2B)
f1.ld.f1(value, time=level, group=Musician, subject=sub,
         time.name="level", group.name="expertise", 
         description=FALSE)$ANOVA.test
```

Main effect of level: F(2.5) = 12.56, p < .001

\newpage
```{r}
pairwise.wilcox.test(data_nested_2B$value, data_nested_2B$level,
                     p.adjust.method = "bonferroni", paired=TRUE)
```
Greater at the 2-bar level than at the 8-bar (p = .045) or 16-bar (p < .001) level, and greater at the 4-bar level than the 16-bar level (p = .0034) (Table S3a). 

\newpage
# Post-hoc tests: comparison of levels (8B)

```{r}
data_nested_8B <- filter(data_nested, scramble == '8B')
```

```{r}
attach(data_nested_8B)
f1.ld.f1(value, time=level, group=Musician, subject=sub,
         time.name="level", group.name="expertise", 
         description=FALSE)$ANOVA.test
```

Main effect of level: F(2.2) = 11.11, p < .001

\newpage
```{r}
pairwise.wilcox.test(data_nested_8B$value, data_nested_8B$level,
                     p.adjust.method = "bonferroni", paired=TRUE)
```

Greater at the 8-bar level than all other levels (2-bar level: p < .001, 4-bar level: p < .001, 16-bar level: p = .0031), and values at the 4-bar level were greater than at the 2-bar level (p < .001) (Table S4a).


\newpage
# Post-hoc tests: comparison of levels (Intact)

```{r}
data_nested_I <- filter(data_nested, scramble == 'Intact')
```

```{r}
attach(data_nested_I)
f1.ld.f1(value, time=level, group=Musician, subject=sub,
         time.name="level", group.name="expertise", 
         description=FALSE)$ANOVA.test
```

No main effect of level: F(2.1) = .830, p = .443

\newpage
```{r}
pairwise.wilcox.test(data_nested_I$value, data_nested_I$level,
                     p.adjust.method = "bonferroni", paired=TRUE)
```
Greater at the 8-bar level than the 4-bar level (p = .0093) (Table S5a).


\newpage
# Interaction between 8B and Intact for 8-bar and 16-bar levels

Look at both groups side by side.

```{r}
long_filtered <- data_nested %>%
  filter(scramble %in% c("8B", "Intact"),
         level %in% c(8,16))

attach(long_filtered)
f1.ld.f2(value, time1=scramble, time2=level, group=Musician, subject=sub,
      time1.name="condition", time2.name="level", group.name="expertise", 
      description=FALSE)$ANOVA.test
```

\newpage
### Musicians only

```{r}
M_long_filtered <- data_nested %>%
  filter(scramble %in% c("8B", "Intact"),
         level %in% c(8,16),
         Musician == "Yes")

attach(M_long_filtered)
ld.f2(value, time1=scramble, time2=level, subject=sub,
      time1.name="condition", time2.name="level", 
      description=FALSE)$ANOVA.test
```
Significant interaction between condition and level for musicians: F(1) = 9.97, p = .0016

\newpage
### Non-musicians 

```{r}
NM_long_filtered <- data_nested %>%
  filter(scramble %in% c("8B", "Intact"),
         level %in% c(8,16),
         Musician == "No")

attach(NM_long_filtered)
ld.f2(value, time1=scramble, time2=level, subject=sub,
      time1.name="condition", time2.name="level", 
      description=FALSE)$ANOVA.test
```

Interaction was not significant for non-musicians: F(1) = 0.331, p = .57