---
title: "Rate"
author: "Riesa Y. Cassano-Coleman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r libraries}
library(tidyverse) # version 2.0.0
library(magrittr) # version 2.0.3
library(nparLD) # version 2.2
select <- dplyr::select # make sure we're using tidyverse's version of select...
```

Load the data.
```{r}
data <- read_csv('../data/response_rate_by_sub.csv', show_col_types = FALSE)
```

Change musician into a factor.
```{r}
data %<>% mutate(Musician = factor(Musician, levels = c('Yes', 'No')))
```

Make sure `Intact` naming is consistent.
```{r}
data %<>% mutate(scramble = ifelse(scramble == 'intact', 'Intact', scramble))
```

\newpage
### Visualize! 
(Generate figure 3, minus significance markers)
```{r}
data %>%
  ggplot(aes(x = scramble, y = mean_response_rate, fill = Musician)) +
  geom_bar(position = "dodge", stat = "summary", fun = mean) +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.2, stat = "summary") +
  geom_point(position = position_jitterdodge(jitter.width = 0.1), alpha = 0.1) +
  theme_gray(base_size = 16) +
  xlab('Scramble Level') +
  ylab('Response Rate (per Minute)') +
  scale_fill_discrete(name="", labels=c('Musicians', 'Non-musicians')) +
  theme(legend.text = element_text(size = 12))
#ggsave('rate.png', width = 7, height = 5)
```

\newpage
### Non-parametric ANOVA-type test

Is there an overall effect of scramble condition or musical experience?
```{r}
attach(data)
# scramble is the within-subject factor ("time") - all subjects hear all conditions
# expertise is the between-subject factor ("group") - subjects are either musicians or non-musicians
f1.ld.f1(mean_response_rate, time=scramble, group=Musician, subject=exp_subject_id,
         time.name="condition", group.name="expertise", description=FALSE)$ANOVA.test
```


* Main effect of scramble condition: F(2.3) = 44.553, p < .001
* No main effect of musical experience: F(1) = 1.407, p = .236
* (Interaction is not significant: F(2.3) = 1.544, p = .210)

\newpage
### Post-hoc tests between conditions

What conditions are different from each other?
```{r}
pairwise.wilcox.test(data$mean_response_rate, 
                     data$scramble, 
                     p.adjust.method = "bonferroni", paired=TRUE)
```

Participants responded more frequently in 1B and 2B compared to 8B and Intact (p < .001 all comparisons). There is no difference between 1B and 2B (p = 1.0) or between 8B and Intact (p = 1.00).
