---
title: "Pitch and Rhythm"
author: "Riesa Y. Cassano-Coleman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r libraries}
library(tidyverse)
library(magrittr)
```

Load the data.
```{r}
data_Intact <- read_csv('pitch_rhythm_resp_Intact.csv', show_col_types = FALSE)
data_8B <- read_csv('pitch_rhythm_resp_8B.csv', show_col_types = FALSE)
data_2B <- read_csv('pitch_rhythm_resp_2B.csv', show_col_types = FALSE)
data_1B <- read_csv('pitch_rhythm_resp_1B.csv', show_col_types = FALSE)
```

\newpage
# Intact

```{r}
summary(lm(n_M_resp ~ abs(pitch_change)*abs(rhythm_change), data_Intact))
```
For musicians, pitch and rhythm explain 0.26% of the variance in number of participants responding.

```{r}
summary(lm(n_NM_resp ~ abs(pitch_change)*abs(rhythm_change), data_Intact))
```
For non-musicians, pitch and rhythm explain 0% of the variance in number of participants responding.

\newpage
# 8B

```{r}
summary(lm(n_M_resp ~ abs(pitch_change)*abs(rhythm_change), data_8B))
```
For musicians, pitch and rhythm explain 4.2% of the variance in number of participants responding.

```{r}
summary(lm(n_NM_resp ~ abs(pitch_change)*abs(rhythm_change), data_8B))
```
For non-musicians, pitch and rhythm explain 2.0% of the variance in number of participants responding.

\newpage
# 2B

```{r}
summary(lm(n_M_resp ~ abs(pitch_change)*abs(rhythm_change), data_2B))
```
For musicians, pitch and rhythm explain 3.1% of the variance in number of participants responding.

```{r}
summary(lm(n_NM_resp ~ abs(pitch_change)*abs(rhythm_change), data_2B))
```
For non-musicians, pitch and rhythm explain 3.4% of the variance in number of participants responding.

\newpage
# 1B

```{r}
summary(lm(n_M_resp ~ abs(pitch_change)*abs(rhythm_change), data_1B))
```
For musicians, pitch and rhythm explain 0.45% of the variance in number of participants responding.

```{r}
summary(lm(n_NM_resp ~ abs(pitch_change)*abs(rhythm_change), data_1B))
```
For non-musicians, pitch and rhythm explain 0% of the variance in number of participants responding.