---
title: "E3 data wrangling"
author: "R. Cassano-Coleman"
date: "`r Sys.Date()`"
output: html_document
---

This notebook combines and wrangles data for experiment 3.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # version 2.0.0
library(magrittr) # version 2.0.3
library(readxl) # version 1.4.3
select <- dplyr::select # make sure we're using the dplyr version of select
```


# Wrangle dataset 3A

Load dataset 3A and the musician/non-musician info.

```{r}
raw_3A <- read_csv('raw_combined_E3A.csv', show_col_types = FALSE)
groups_3A <- read_excel('../E3/sub_ids_3A.xlsx')
addl_subs <- read_excel('subs_additional_E3A.xlsx')
```

Keep only the main task.

```{r}
main_3A <- raw_3A %>%
  filter(grepl('OpenEndedSeg',Task_Name)) %>%
  # scramble level is in one of the following columns...
  mutate(scramble = ifelse(!is.na(scramble_cat_4), scramble_cat_4, 
                           ifelse(!is.na(scramble_level), scramble_level,
                                  ifelse(!is.na(scramble_open_3), scramble_open_3, NA)))) %>%
  select(c(exp_subject_id, Trial_Nr, Task_Name, participant_spacePress2, scramble)) %>%


  # add group info
  left_join(., groups_3A, by = join_by('exp_subject_id')) %>%
  relocate(Musician, .after = exp_subject_id) %>%
  # subjects who are left out did stimulus set 2, which is highly overlapping with stimulus set 1
  filter(!is.na(Musician))
```

Add the four additional subjects.

```{r}
main_3A <- addl_subs %>%
  select(colnames(main_3A)) %>%
  bind_rows(main_3A, .) %>%
  # capitalize Intact
  mutate(scramble = ifelse(scramble == 'intact', 'Intact', scramble))
```

Additional wrangling...
  
```{r}
main_3A %<>%
  # the number in Task_Name encodes stimulus set
  separate_wider_delim(Task_Name, delim = '_', names_sep = '_') %>% #, names = c(NULL, 'stimulus_set')) # not sure why this didn't work
  mutate(Task_Name_1 = NULL) %>%
  rename(stimulus_set = Task_Name_2) %>%

  # keep only the actual timestamps from `space_press2`
  # can't specify the column names in this step because there are some cells with NAs
  separate_wider_delim(participant_spacePress2, delim = '---', names_sep = '_') %>%
  mutate(participant_spacePress2_1 = NULL) %>%
  rename(responses = participant_spacePress2_2) %>%

  # transform response column into timestamp representation
  # first, remove NA rows
  filter(!is.na(responses)) %>%
  # replace 'values=' and " with an empty string
  mutate(responses = str_replace(responses, 'values=', '')) %>%
  mutate(responses = str_replace(responses, '"', '')) %>%
  # give each timestamp its own row
  separate_longer_delim(responses, delim = ';') %>%
  filter(responses != '') %>%
  # convert from milliseconds to seconds
  mutate(responses = as.numeric(responses) / 1000) %>%

  # rename column for analysis
  rename(value = responses) %>%
  
  # move scramble
  relocate(scramble, .after = stimulus_set)
```


Compute rate.

```{r}
rate_3A <- main_3A %>%
  group_by(exp_subject_id, Musician, stimulus_set, scramble, Trial_Nr) %>%
  summarize(rate = (n() / 64) * 60)
```

Check for rates > 30

```{r}
rate_3A %>%
  filter(rate > 30)
```

Average over all trials in a condition.
```{r}
rate_3A %<>%
  group_by(exp_subject_id, Musician, stimulus_set, scramble) %>%
  summarize(mean_response_rate = mean(rate)) %>%
  filter(mean_response_rate < 30)
```



# Wrangle dataset 3B

Load dataset 3B.

```{r}
raw_3B <- read_csv('raw_combined_E3B.csv', show_col_types = FALSE)
```

Select necessary columns and keep only main task.

```{r}
main_3B <- raw_3B %>%
  # filter the main task
  # Block_Nr = 1 is before the main task: consent, instructions, headphone check, practice
  # Block_Nr = 3 is the post-survey
  filter(Block_Nr == '2') %>%
  
  # keep only the necessary columns
  # just want the responses from run 2, but add 'space_press1' here if you want to look at run 1 as well
  select('Task_Name', 'space_press2', 'exp_subject_id') %>%
  # move subject ID to the front
  relocate('exp_subject_id') %>%
  
  # any task name with a percent in it is a "strategy" question
  filter(!grepl('%', Task_Name)) %>%

  # separate condition and stimulus number
  separate_wider_delim(Task_Name, delim = '_', names = c('scramble', 'stim_num')) %>%
  # capitalize Intact
  mutate(scramble = ifelse(scramble == 'intact', 'Intact', scramble)) %>%
  
  # keep only the actual timestamps from `space_press2`
  # can't specify the column names in this step because there are some cells with NAs
  separate_wider_delim(space_press2, delim = '---', names_sep = '_') %>%
  mutate(space_press2_1 = NULL) %>%
  rename(responses = space_press2_2) #%>%
  
  # before cleaning up this response column, check that each subject has complete data
  #group_by(exp_subject_id) %>%
  #summarize(count = n())
  # all of the subject have 12 trials (complete), except for 377660 who isn't included in the subsetting later anyway
```


Subset dataset 3B to match musician criteria and stimulus length of dataset 3A.

```{r}
# list of subject IDs that match the musician criteria (either >5 years of experience or none at all)
# musicians (n = 13)
mus_sub_ids <- c(
  '377647', '393320', '393267', '393245', '393246',
  '393242', '393241', '393240', '393230', '393254', 
  '393252', '393249', '393239')

# non-musicians (n = 10)
non_mus_sub_ids <- c(
  '377777', '377770', '377747', '377708', '377701',
  '377692', '377665', '377664', '377663', '393253')

subset_3B <- main_3B %>%
  # musician criteria
  mutate(Musician = ifelse(exp_subject_id %in% mus_sub_ids, 'Yes',
                           ifelse(exp_subject_id %in% non_mus_sub_ids, 'No', NA))) %>%
  relocate(Musician, .after = exp_subject_id) %>%
  filter(!is.na(Musician)) %>%
  
  # stimulus length
  
  # remove stimulus 2
  filter(stim_num != '2') %>%

  # transform response column into timestamp representation
  # first, remove NA rows
  filter(!is.na(responses)) %>%
  # replace 'values=' and " with an empty string
  mutate(responses = str_replace(responses, 'values=', '')) %>%
  mutate(responses = str_replace(responses, '"', '')) %>%
  # give each timestamp its own row
  separate_longer_delim(responses, delim = ';') %>%
  filter(responses != '') %>%
  # convert from milliseconds to seconds
  mutate(responses = as.numeric(responses) / 1000) %>%

  # subsetting rules:
  # when stim_num == 1, take all timestamps under 66 seconds
  filter(!(stim_num == '1' & responses > 66)) %>%
  # when stim_num == 3, take the last 66 seconds
  # everything after 16 seconds for Intact (since Intact is 41m/82s)
  filter(!(stim_num == '3' & scramble == 'Intact' & responses < 16)) %>%
  # everything after 10 seconds for Intact (since 8B is 38m/76s)
  filter(!(stim_num == '3' & scramble == '8B' & responses < 10)) %>%
  # everything after 14 seconds for 2B (since 2B is 40m/80s)
  filter(!(stim_num == '3' & scramble == '2B' & responses < 14)) %>%
  # everything after 14 seconds for 1B (since 1B is 40m/80s)
  filter(!(stim_num == '3' & scramble == '1B' & responses < 14)) %>%

  # adjust timing for stim_num == 3 so it starts at 0 for all conditions  
  mutate(responses = ifelse(stim_num == 3, ifelse(
    scramble == 'Intact', responses - 16, ifelse(
      scramble == '8B', responses - 10, responses - 14
      )), responses)) %>%
  # check that nothing is below zero
  #filter(responses < 0)
  
  # rename column for analysis
  rename(value = responses) %>%
  
  # add stimulus set = 1 for when we combine
  mutate(stimulus_set = 1, .after = Musician)
```

Compute rate and exclude any cases where rate is greater than 1 response every 2 seconds (30 responses per minute).

```{r}
# each of these stimuli are 66 seconds, rate is expressed per minute
rate_3B <- subset_3B %>% 
  group_by(exp_subject_id, Musician, stimulus_set, scramble, stim_num) %>%
  summarize(rate = (n() / 66) * 60) %>%
  # average over both runs for each subject
  group_by(exp_subject_id, Musician, stimulus_set, scramble) %>%
  summarize(mean_response_rate = mean(rate))
```

None of these cases have rates > 30.


# Combine

## Rate

```{r}
rate_3A$stimulus_set <- as.numeric(rate_3A$stimulus_set) # to match 3B
rate_combined <- bind_rows(rate_3A, rate_3B)
```

Check for any NAs
```{r}
rate_combined %>% group_by(exp_subject_id) %>% summarize(count = n())
```

Missing data: 
- has data for 1 condition: 355341, 356778, 361615, 363111 (12)
- has data for 2 conditions: 354009, 360247 (4)
- has data for 3 conditions: 334739, 360430, 385662, 393253, 610499 (5)
This explains the difference between the reference (380 rows) and what I just got (359 rows). Need to compute the means without dropping NAs.


Make things into factors so we don't have to do it later.

```{r}
rate_combined %<>%
  mutate(subexp_subject_id = factor(exp_subject_id),
         Musician = factor(Musician, levels = c('Yes', 'No')),
         stimulus_set = factor(stimulus_set),
         scramble = factor(scramble, levels = c('Intact', '8B', '2B', '1B'))
         )
```


## Timestamps

subset_3B

