---
title: "E3 rate"
author: "R. Cassano-Coleman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
knit: (function(inputFile, encoding) {rmarkdown::render(inputFile, encoding = encoding, output_dir = "notebooks") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # version 2.0.0
library(magrittr) # version 2.0.3
library(rstatix) # version 0.7.2

library(brms) # version 2.22.0
library(bayestestR) # version 0.15.0
library(emmeans) # version 1.10.0
```

This notebook takes response rates by subject, log-transforms them, and runs a Bayesian version of a mixed effects model.

```{r}
set.seed(15000)
```

Load the data.

```{r}
data <- read_csv('../data/E3/response_rate_by_sub.csv', show_col_types = FALSE)
```

Change musician and scramble into a factor.

```{r}
data %<>% mutate(Musician = factor(Musician, levels = c('Yes', 'No')),
                 scramble = factor(scramble, levels = c('Intact', '8B', '2B', '1B')))
```

Set Intact as reference level.

```{r}
contrasts(data$scramble) <- contr.treatment(4)
```

\newpage

Check normality of the true response rate data. (Ignoring stimulus set.)

```{r}
data %>% 
  group_by(Musician, scramble) %>%
  shapiro_test(mean_response_rate)
```

```{r}
data %>%
  ggplot(aes(x = scramble, y = mean_response_rate, fill = Musician)) +
  geom_bar(position = "dodge", stat = "summary", fun = mean) +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.2, stat = "summary") +
  geom_point(position = position_jitterdodge(jitter.width = 0.1), alpha = 0.1) +
  theme_gray(base_size = 16) +
  xlab('Scramble Level') +
  ylab('Response Rate (per Minute)') +
  scale_fill_discrete(name="", labels=c('Musicians', 'Non-musicians')) +
  theme(legend.text = element_text(size = 12))
```

\newpage

Log-transform the rates and check for normality.

```{r}
data %<>% mutate(log_rate = log(1 + mean_response_rate))
# add 1 so rates that are zero transform to 0 (rather than negative infinity)
```

```{r}
data %>% 
  group_by(Musician, scramble) %>%
  shapiro_test(log_rate)
```

Some of these are worse than others. Visualize:

```{r}
data %>%
  ggplot(aes(x = scramble, y = log_rate, fill = Musician)) +
  geom_bar(position = "dodge", stat = "summary", fun = mean) +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.2, stat = "summary") +
  geom_point(position = position_jitterdodge(jitter.width = 0.1), alpha = 0.1) +
  theme_gray(base_size = 16) +
  xlab('Scramble Level') +
  ylab('Log of Response Rate (per Minute)') +
  scale_fill_discrete(name="", labels=c('Musicians', 'Non-musicians')) +
  theme(legend.text = element_text(size = 12))
#ggsave('log_rate.png', width = 7, height = 5)
```

It seems like this lack of normality is driven by the zero rates.



\newpage

```{r}
get_prior(log_rate ~ Musician + scramble + (1|exp_subject_id), data = data)
```

```{r}
these_priors <- c(
  set_prior('normal(0, 0.5)', coef = "MusicianNo"), # don't necessarily expect a difference between groups
  set_prior('normal(0, 0.5)', coef = "scramble2"), # intact vs 8B 
  set_prior('normal(0, 0.5)', coef = "scramble3"), # intact vs 2B 
  set_prior('normal(0, 0.5)', coef = "scramble4") # intact vs 1B 
)
```

```{r}
brm_log_rate <- brm(log_rate ~ Musician + scramble + (1|exp_subject_id), data = data,
                    prior = these_priors,
                    save_pars = save_pars(all = TRUE), 
                    iter = 20000, refresh = 0,
                    file = 'models/E3_log_rate')
```

\newpage

```{r}
plot(brm_log_rate)
print(summary(brm_log_rate), digits = 4)
```

\newpage

```{r}
emm_log_rate <- emmeans(brm_log_rate, specs = c("scramble", "Musician"))
summary(emm_log_rate)
```

```{r}
emm_log_rate_s <- emmeans(brm_log_rate, specs = "scramble")
summary(emm_log_rate_s)
```

```{r}
contrast(emm_log_rate_s, method = "pairwise")
```

\newpage

```{r}
log_rate_BF <- describe_posterior(brm_log_rate,
                                  estimate = "median", dispersion = TRUE,
                                  ci = .95, ci_method = "HDI",
                                  test = c("bayes_factor"))
print(log_rate_BF, digits = 4)
```

\newpage

Compare the full model to a model without scramble condition.

```{r}
brm_log_rate_null <- brm(log_rate ~ Musician + (1|exp_subject_id), data = data,
                         prior = set_prior('normal(0, 0.5)', class = 'b'),
                         save_pars = save_pars(all = TRUE), 
                         iter = 20000, refresh = 0,
                         file = 'models/E3_log_rate_null')
```

```{r}
plot(brm_log_rate_null)
print(summary(brm_log_rate_null), digits = 4)
```

```{r}
BF_log_rate <- bayes_factor(brm_log_rate, brm_log_rate_null)
print(BF_log_rate)
```

\newpage

Visualize with posterior estimates and 95% CrI 

```{r}
posterior_est <- as.data.frame(emm_log_rate)
```

```{r}
ggplot() +
  geom_col(aes(x = scramble, y = exp(emmean), fill = Musician), data = posterior_est,
           position = "dodge") +
  geom_errorbar(aes(x = scramble, ymin = exp(lower.HPD), ymax = exp(upper.HPD), fill = Musician), 
                data = posterior_est, position = position_dodge(width = 0.9), width = 0.2) +
  geom_point(aes(x = scramble, y = mean_response_rate, fill = Musician), data = data, 
             position = position_jitterdodge(dodge.width = 0.9, jitter.width = 0.1), alpha = 0.1) +
  theme_gray(base_size = 16) +
  scale_x_discrete(limits = rev) +
  xlab('Scramble Level') +
  ylab('Response Rate (per Minute)') +
  scale_fill_discrete(name="", labels=c('Musicians', 'Non-musicians')) +
  theme(legend.text = element_text(size = 12))

ggsave('../figures/Fig3_rate.png', width = 7, height = 5)
```
